The controlled choice design and private paternalism in pawnshop borrowing - STATA documentation

### 2024-05-24 Restud submission/donde_restud.tex
new submission

% use option [draft] for initial submission
%            [final] for the prepublication
\documentclass[ecta,nameyear,draft]{econsocart}
%\documentclass[]{econsocart}
%
%\usepackage{}
\RequirePackage[colorlinks,citecolor=blue,linkcolor=blue,urlcolor=blue,pagebackref]{hyperref}

\startlocaldefs

\usepackage{booktabs}
\usepackage{multirow}
\usepackage{float}         
\usepackage{amsmath}
\usepackage{amscd}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{bbm}
\usepackage{dsfont}
\usepackage{colortbl}
\usepackage{comment}
\usepackage{lscape}
\usepackage{subcaption}
\newtheorem{theorem}{Theorem}
\newtheorem{claim}[theorem]{Claim}
\newtheorem{prop}[theorem]{Proposition} 
\newtheorem{cor}[theorem]{Corollary} 
\newtheorem{assumption}{Assumption} 
\newtheorem{lem}{Lemma} 
\usepackage{enumerate}

\newcommand{\wh}{\widehat}

\newcommand{\E}{\mathbb{E}}

\endlocaldefs

\usepackage[section]{placeins}

\begin{document}

\begin{frontmatter}

\title{The controlled choice design and private paternalism in pawnshop borrowing}
\runtitle{The controlled choice design and private paternalism}

\begin{aug}
% use \particle for den|der|de|van|von (only lc!)
% [id=?,addressref=?,corref]{\fnms{}~\snm{}\ead[label=e?]{}\thanksref{}}
%
%% e-mail is mandatory for each author
%
%%% initials in fnms (if any) with spaces
%
\author[id=au1,addressref={add1}]{\fnms{Craig}~\snm{McIntosh}\ead[label=e1]{ctmcintosh@ucsd.edu}}
\author[id=au2,addressref={add2}]{\fnms{Isaac}~\snm{Meza}\ead[label=e2]{isaacmezalopez@g.harvard.edu}}
\author[id=au3,addressref={add3}]{\fnms{Joyce}~\snm{Sadka}\ead[label=e3]{jsadka@itam.mx}}
\author[id=au4,addressref={add4}]{\fnms{Enrique}~\snm{Seira}\ead[label=e4]{enrique.seira@gmail.com}}
\author[id=au5,addressref={add5}]{\fnms{Francis J.}~\snm{DiTraglia}\ead[label=e5]{francis.ditraglia@economics.ox.ac.uk}}

% Addresses
\address[id=add1]{%
\orgdiv{Department of Economics}
\orgname{University of California San Diego}}

\address[id=add2]{%
\orgdiv{Department of Economics}
\orgname{Harvard University}}

\address[id=add3]{%
\orgdiv{Department of Economics}
\orgname{Instituto Tecnologico Autonomo de Mexico}}

\address[id=add4]{%
\orgdiv{Department of Economics}
\orgname{Michigan State University}}

\address[id=add5]{%
\orgdiv{Department of Economics}
\orgname{University of Oxford}}
\end{aug}

%% Put support info here. Reminder: do not thank the handling coeditor anonymously or by name
\support{We want to thank Mauricio Romero and Anett John for advice and encouragement. Ricardo Olivares, Gerardo Melendez, and Alonso de Gortari provided excellent research assistance and Erick Molina helped with formatting. Jose Maria Barrero, Andrei Gomberg, Emilio Gutierrez, David Laibson, Aprajit Mahajan, Matt Rabin, Charlie Sprenger, and seminar participants at ITAM, USC, MSU, and UCSD provided valuable feedback. Research assistance was financed through faculty grants at ITAM. Our research partner had no say in the results.}
%
\coeditor{\fnm{[Name} \snm{Surname}; will be inserted later]}

@@

\begin{keyword}
\kwd{Private paternalism}
\kwd{choice}
\kwd{treatment on the untreated}
\kwd{heterogeneous treatment effects}
\kwd{commitment}
\kwd{overconfidence}
\end{keyword}

\end{frontmatter}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%% Main text entry area:
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

As a second measure of cost we calculate the Annual Percentage Rate (APR) in order to express the cost as a percentage of the loan, per year, inclusive of default costs. The standard definition is given in the following formula: $(\text{APR})_i =\left( 1 + \frac{\frac{\text{Financial Cost}_i}{\text{Loan}_i}}{\text{loan term}_i}\right)^{\text{loan term}_i}-1 $

@@

\begin{figure}[htbp]    
\begin{center}
\includegraphics[width=0.9\textwidth]{Figuras/consort.pdf}
  \end{center}
 \caption{Experiment description}
     \label{exp_description}
%\textit{Do file: }  \texttt{consort.do}
\end{figure}

@@

\begin{table}
\caption{Effects on Financial Cost}
\label{main_impact_table}
\begin{center}
\resizebox{1.0\textwidth}{!}{
\scriptsize{\input{./Tables/decomposition_main_te.tex}}
}
\end{center}
\legend{This table shows the treatment effects for our core pecuniary outcomes. Each column is a different regression for different outcomes on an indicator for the forced and choice arms, following specification in equation \ref{basic_reg}. Columns (1) \& (7) analyze our core financial cost measures, while the rest of the columns decompose these into finer components.
A few borrowers take more than one loan on the first day they appear in an experimental branch. These are treated as different observations. Additional results, available upon request, show that our results are robust to different ways of handling multiple loans for each borrower. Each regression includes branch and day-of-week FE. Standard errors are clustered at the branch-day level. }
%Table \ref{multiple_loans} shows our results are similar when dealing with the multiplicity of loans per borrower.

@@

\begin{table}
\caption{Effects on Repeat Pawning}
\label{repeat_loans}
\begin{center}
\scriptsize{\input{./Tables/repeat_loans.tex}}
\end{center}
 \legend{This table estimates the specification of equation \ref{basic_reg} but at the level of the borrower (not the loan). Each column represents a regression with a different outcome variable.
 % In column 1, the dependent variable indicates, for each borrower in the experiment, whether he/she pawned again after the first loan in the experiment (up to the end period of our data set 338 days after the experiment began). Column 2 is analogous, but only pawning after 90 days of the first loan is considered. Column 3 instead considers pawning before 90 days. Column 4 is analogous to column 1 but focuses on the pawning of a gold piece that is different from the one in the first experimental loan. Column 5 is analogous to column 1, but conditioning on the sample that recovered the first loan. 
 Each regression includes branch and day-of-week FE. Standard errors are clustered at the branch-day level.}
%\textit{Do file: } \texttt{repeat\_loans.do}
\end{table}

@@

\begin{figure}
    \begin{center}
        \centering
        \includegraphics[width=1.0\textwidth]{Figuras/tot_tut_intuition.png}
    \end{center}
 \caption{Graphical Intuition for the Controlled Choice Design. 
     %This figure illustrates why the controlled choice design identifies the TOT and TUT effects.
    Gray rectangles denote borrowers with a commitment contract; white rectangles denote borrowers with a status quo contract. 
    A comparison of means across control and forcing arms identifies the ATE of forcing commitment. 
    %In the choice arm, anyone with a commitment contract is a chooser; in the control and forcing arms, we do not know who is a chooser and who is not, but because borrowers are randomly assigned to experimental treatment arms, the share of choosers will be the same, on average (depicted using dashed vertical lines). 
    The difference of mean outcomes across the choice and control ``nets out'' the non-choosers, and hence equals the TOT multiplied by the share of choosers.
    %arms gives the intent-to-treat effect of offering choice. 
    %Under the exclusion restriction that moving non-choosers between the choice and forcing arms does not change their outcomes, this comparison ``nets out'' the non-choosers.
    %Hence, the ITT of offering choice equals the TOT multiplied by the share of choosers. 
    %Similarly, under the exclusion restriction that moving choosers between the choice and forcing arm does not affect their outcomes. 
    Similarly, the difference of means across the forcing and choice arms ``nets out'' the choosers and equals the TUT multiplied by the share of non-choosers.
        The share of choosers, illustrated using dashed vertical lines, is equal on average across arms under random assignment.}
    \label{tot_tut_graph}
     %\textit{Do file: }  \texttt{tot\_tut\_graph.do}
\end{figure}

@@

\begin{table}[H]
\caption{Treatment on the Treated (TOT), Treatment on the Untreated (TUT), Selection-on-gains (TOT - TUT), Average Selection Bias (ASB), and Average Selection Bias.}
\label{tot_tut}
\begin{center}
\scriptsize{\input{./Tables/tot_tut.tex}}
\end{center}
\legend{This table presents results computed using the derivations from Section \ref{sec:randchoice}. The APR and financial cost outcomes have been multiplied by $-1$ so that a positive causal effect \emph{benefits} the borrower in each of the four columns. The bottom panel present p-values for a number of null hypothesis tests of treatment effect heterogeneity. }
%\textit{Do file: } \texttt{tot\_tut.do}
\end{table}

@@

\begin{figure}[H]
\begin{center}
        \centering
        \includegraphics[width=0.6\textwidth]{Figuras/line_cw_apr_tot_tut.pdf}

 \end{center}       
 \caption{``Mistakes'' in the choice arm. This figure presents conditional average TUT and TOT effects for the APR outcome from Figure \ref{heterogeneous_effects} in an alternative manner, to consider the fraction of borrowers in the choice arm who made ``mistakes'' in their decision to accept or refuse the commitment contract. A ``mistake'' for a non-chooser is defined as a positive conditional TUT effect that significantly exceeds a specified threshold APR value.
%  The green curve equals $[1 - F_{\text{TUT}}(\delta)]\times 100\%$, where $F_{\text{TUT}}$ is the CDF corresponding to density of conditional TUT effects from Figure \ref{heterogeneous_effects}, computed using the instrumental forest approach from \cite{atheygrf}. Evaluated at any positive value on the horizontal axis, it gives the fraction of non-choosers in the choice arm who made a ``mistake'' by not choosing commitment. The green shaded region gives associated 95\% pointwise confidence bands. 
% Analogously, a ``mistake'' for a chooser is defined as a negative conditional TOT effect that exceeds a specified threshold APR value. The red curve equals $F_{\text{TOT}}(-\delta)$ where $F_{\text{TOT}}$ is the CDF corresponding to the density of conditional TOT effects from Figure \ref{heterogeneous_effects}. 
% The red shaded region gives associated 95\% pointwise confidence bands.
% For both the red and green curves, we define the APR threshold so that larger mistakes are to the \emph{right} of smaller mistakes.
% This allows us to construct the overall fraction of mistakes in the choice arm, the blue curve, as a weighted average of the green and red curves.
% The weights equal the share of choosers and non-choosers in the choice arm.
}
    \label{choose_wrong}
\end{figure}

@@

\begin{table}
\caption{Type I \& II errors using targeting narrow rules}
\label{hit_miss_rule}
\begin{center}

\scriptsize{\input{./Tables/hit_miss_rule.tex}}

\end{center}
\legend{This table reports error rates for six different rules for allocating individuals to commitment. Row 1 assigns all borrowers to control, Row 2 all to the Forcing arm. Row 3 uses `optimal targeting according to the CATE from the wide covariate set. Row 4 uses a random forest classification, and Row 5 a logit model, both with only the narrow covariate set.  Row 6 uses the choices made by borrowers to assign to commitment.}
\end{table}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%BIBLIOGRAPHY

\bibliographystyle{ecta-fullname} % Style BST file
\bibliography{References}  % Bibliography file (usually '*.bib')

@@

\input{ECTA submission/OA_ecta}
\end{document}

>>>>

\documentclass[12pt, a4paper]{article}

\input{preamble.tex}
%%% HELPER CODE FOR DEALING WITH EXTERNAL REFERENCES
\usepackage{xr}
\makeatletter
\newcommand*{\addFileDependency}[1]{
  \typeout{(#1)}
  \@addtofilelist{#1}
  \IfFileExists{#1}{}{\typeout{No file #1.}}
}
\makeatother

\newcommand*{\myexternaldocument}[1]{
    \externaldocument{#1}
    \addFileDependency{#1.tex}
    \addFileDependency{#1.aux}
}

\newgeometry{margin=1in}
\linespread{1.25}
\begin{document}

\title{The controlled choice design and private paternalism in pawnshop borrowing\thanks{We want to thank Mauricio Romero and Anett John for advice and encouragement. Ricardo Olivares, Gerardo Melendez, and Alonso de Gortari provided excellent research assistance and Erick Molina helped with formatting. Jose Maria Barrero, Andrei Gomberg, Emilio Gutierrez, David Laibson, Aprajit Mahajan, Matt Rabin, Charlie Sprenger, and seminar participants at ITAM, USC, MSU, and UCSD provided valuable feedback. Research assistance was financed through faculty grants at ITAM. Our research partner had no say in the results.}}
\author{Craig McIntosh \and Isaac Meza \and Joyce Sadka \and Enrique Seira \and Francis J.\ DiTraglia   \thanks{Seira:  MSU, \url{enrique.seira@gmail.com} (corresponding author); McIntosh:  University of California San Diego, \url{ctmcintosh@ucsd.edu}; Meza: Harvard University, \url{isaacmezalopez@g.harvard.edu}; Sadka: ITAM, \url{jsadka@itam.mx}; DiTraglia: Oxford, \url{francis.ditraglia@economics.ox.ac.uk}} }
\date{This draft:  \today \\[2 cm]}

%\vspace{.5in}

\maketitle
\vspace{-0.75in}

@@

\textbf{Keywords: } Private paternalism, choice, treatment on the untreated, heterogeneous treatment effects, commitment, overconfidence.

\textbf{JEL codes:} G41, C93, O16, G21

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%% Main text entry area:
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

As a second measure of cost we calculate the Annual Percentage Rate (APR) in order to express the cost as a percentage of the loan, per year, inclusive of default costs. The standard definition is given in the following formula: 
$$(\text{APR})_i =\left( 1 + \frac{\frac{\text{Financial Cost}_i}{\text{Loan}_i}}{\text{loan term}_i}\right)^{\text{loan term}_i}-1 $$

@@

\begin{figure}[htbp]    
 \caption{Experiment description}
\begin{center}
\includegraphics[width=0.9\textwidth]{Figuras/consort.pdf}
  \end{center}
     \label{exp_description}
%\textit{Do file: }  \texttt{consort.do}
\end{figure}

@@

\begin{table}
\caption{Effects on Financial Cost}
\label{main_impact_table}
\begin{center}
\resizebox{1.0\textwidth}{!}{
\footnotesize{\input{./Tables/decomposition_main_te.tex}}
}
\end{center}
\footnotesize{This table shows the treatment effects for our core pecuniary outcomes. Each column is a different regression for different outcomes on an indicator for the forced and choice arms, following specification in equation \ref{basic_reg}. Columns (1) \& (7) analyze our core financial cost measures, while the rest of the columns decompose these into finer components.
A few borrowers take more than one loan on the first day they appear in an experimental branch. These are treated as different observations. Additional results, available upon request, show that our results are robust to different ways of handling multiple loans for each borrower. Each regression includes branch and day-of-week FE. Standard errors are clustered at the branch-day level. }
%Table \ref{multiple_loans} shows our results are similar when dealing with the multiplicity of loans per borrower.

@@

\begin{table}
\caption{Effects on Repeat Pawning}
\label{repeat_loans}
\begin{center}
\footnotesize{\input{./Tables/repeat_loans.tex}}
\end{center}
 \footnotesize{This table estimates the specification of equation \ref{basic_reg} but at the level of the borrower (not the loan). Each column represents a regression with a different outcome variable.
 % In column 1, the dependent variable indicates, for each borrower in the experiment, whether he/she pawned again after the first loan in the experiment (up to the end period of our data set 338 days after the experiment began). Column 2 is analogous, but only pawning after 90 days of the first loan is considered. Column 3 instead considers pawning before 90 days. Column 4 is analogous to column 1 but focuses on the pawning of a gold piece that is different from the one in the first experimental loan. Column 5 is analogous to column 1, but conditioning on the sample that recovered the first loan. 
 Each regression includes branch and day-of-week FE. Standard errors are clustered at the branch-day level.}
%\textit{Do file: } \texttt{repeat\_loans.do}
\end{table}

@@

\begin{figure}
\caption{Graphical Intuition for the Controlled Choice Design.} 
    \begin{center}
        \centering
        \includegraphics[width=1.0\textwidth]{Figuras/tot_tut_intuition.png}
    \end{center}
 \footnotesize{
     %This figure illustrates why the controlled choice design identifies the TOT and TUT effects.
    Gray rectangles denote borrowers with a commitment contract; white rectangles denote borrowers with a status quo contract. 
    A comparison of means across control and forcing arms identifies the ATE of forcing commitment. 
    %In the choice arm, anyone with a commitment contract is a chooser; in the control and forcing arms, we do not know who is a chooser and who is not, but because borrowers are randomly assigned to experimental treatment arms, the share of choosers will be the same, on average (depicted using dashed vertical lines). 
    The difference of mean outcomes across the choice and control ``nets out'' the non-choosers, and hence equals the TOT multiplied by the share of choosers.
    %arms gives the intent-to-treat effect of offering choice. 
    %Under the exclusion restriction that moving non-choosers between the choice and forcing arms does not change their outcomes, this comparison ``nets out'' the non-choosers.
    %Hence, the ITT of offering choice equals the TOT multiplied by the share of choosers. 
    %Similarly, under the exclusion restriction that moving choosers between the choice and forcing arm does not affect their outcomes. 
    Similarly, the difference of means across the forcing and choice arms ``nets out'' the choosers and equals the TUT multiplied by the share of non-choosers.
        The share of choosers, illustrated using dashed vertical lines, is equal on average across arms under random assignment.}
    \label{tot_tut_graph}
     %\textit{Do file: }  \texttt{tot\_tut\_graph.do}
\end{figure}

@@

\begin{table}[H]
\caption{Treatment on the Treated (TOT), Treatment on the Untreated (TUT), Selection-on-gains (TOT - TUT), Average Selection Bias (ASB), and Average Selection Bias.}
\label{tot_tut}
\begin{center}
\footnotesize{\input{./Tables/tot_tut.tex}}
\end{center}
\footnotesize{This table presents results computed using the derivations from Section \ref{sec:randchoice}. The APR and financial cost outcomes have been multiplied by $-1$ so that a positive causal effect \emph{benefits} the borrower in each of the four columns. The bottom panel present p-values for a number of null hypothesis tests of treatment effect heterogeneity. }
%\textit{Do file: } \texttt{tot\_tut.do}
\end{table}

@@

\begin{figure}[H]
 \caption{``Mistakes'' in the choice arm.}
\begin{center}
        \centering
        \includegraphics[width=0.65\textwidth]{Figuras/line_cw_apr_tot_tut.pdf}
 \end{center}       
 \footnotesize{This figure presents conditional average TUT and TOT effects for the APR outcome from Figure \ref{heterogeneous_effects} in an alternative manner, to consider the fraction of borrowers in the choice arm who made ``mistakes'' in their decision to accept or refuse the commitment contract. A ``mistake'' for a non-chooser is defined as a positive conditional TUT effect that significantly exceeds a specified threshold APR value.
%  The green curve equals $[1 - F_{\text{TUT}}(\delta)]\times 100\%$, where $F_{\text{TUT}}$ is the CDF corresponding to density of conditional TUT effects from Figure \ref{heterogeneous_effects}, computed using the instrumental forest approach from \cite{atheygrf}. Evaluated at any positive value on the horizontal axis, it gives the fraction of non-choosers in the choice arm who made a ``mistake'' by not choosing commitment. The green shaded region gives associated 95\% pointwise confidence bands. 
% Analogously, a ``mistake'' for a chooser is defined as a negative conditional TOT effect that exceeds a specified threshold APR value. The red curve equals $F_{\text{TOT}}(-\delta)$ where $F_{\text{TOT}}$ is the CDF corresponding to the density of conditional TOT effects from Figure \ref{heterogeneous_effects}. 
% The red shaded region gives associated 95\% pointwise confidence bands.
% For both the red and green curves, we define the APR threshold so that larger mistakes are to the \emph{right} of smaller mistakes.
% This allows us to construct the overall fraction of mistakes in the choice arm, the blue curve, as a weighted average of the green and red curves.
% The weights equal the share of choosers and non-choosers in the choice arm.
}
    \label{choose_wrong}
\end{figure}

@@

\begin{table}
\caption{Type I \& II errors using targeting narrow rules}
\label{hit_miss_rule}
\begin{center}
\footnotesize{\input{./Tables/hit_miss_rule.tex}}
\end{center}
\footnotesize{This table reports error rates for six different rules for allocating individuals to commitment. Row 1 assigns all borrowers to control, Row 2 all to the Forcing arm. Row 3 uses `optimal targeting according to the CATE from the wide covariate set. Row 4 uses a random forest classification, and Row 5 a logit model, both with only the narrow covariate set.  Row 6 uses the choices made by borrowers to assign to commitment.}
\end{table}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%BIBLIOGRAPHY
\begingroup
\setstretch{1.10}
\bibliographystyle{authordate1}
%\bibliographystyle{amsalpha}
%\bibliographystyle{AER}

\bibliography{References}
\endgroup

@@

\input{Restud submission/OA_restud}
\end{document}
